### 2015-02-28 icfp15-servo/servo.tex
Expand on compilation strategy

\section{Servo}
\label{sec:servo}
A crucial test of Servo is performance --- Servo must be at least as fast
as other browsers at similar tasks to succeed, even if it provides additional memory safety.
\tblref{servo-perf} shows a preliminary comparison of the performance of the layout stage (described
in \secref{sec:browsers}) of rendering several web sites in Mozilla Firefox's Gecko engine compared
to Servo, taken on a modern MacBook Pro.
\begin{table}
  \begin{center}
    \begin{tabular}{r || r | r | r}
      Site & Gecko & Servo 1 thread & Servo 4 threads \\
      \hline
      Reddit & 250 & 100 & 55  \\
      CNN & 105 & 50 & 35 \\
    \end{tabular}%
  \end{center}%
  \caption{Performance of Servo against Mozilla's Gecko rendering engine on the layout portion of some common sites.
  Times are in milliseconds, where lower numbers are better.}
  \label{servo-perf}
\end{table}
In the remainder of this section, we cover specific areas of Servo's design or implementation that make use of
Rust and the impacts and limitations of these features.

@@

\subsection{Compilation strategy}
Many statically typed implementations of polymorphic languages such as Standard ML of New Jersey~\cite{SMLNJ} and
\ocaml{}~\cite{ocaml-manual-3.0} have used a compilation strategy that optimizes representations of data types when
polymorphic code is monomorphically used, but defaults to a less efficient style otherwise, in order to share
code~\cite{ocaml-repr}.
This strategy reduces code size, but leads to unpredictable performance and code, as changes to a codebase that
either add a new instantiation of a polymorphic function at a given type or, in a modular compilation setting, that
expose a polymorphic function externally, can change the performance of code that is not local to the change being
made.
Monomorphization, as in MLton~\cite{mlton-compiler}, instead instantiates each polymorphic code block at each of the types
it is applied against, providing predictable output code to developers at the cost of some code duplication.
This approach, combined with whole-program compilation~\cite{weeks:whole-program-mlton}, is used in Rust.
This strategy has a negative impact on compilation speed but has resulted in Rust code that easily matches the
sequential speed of its \Cplusplus{} analog without requiring the Servo developers to become compiler experts.

>>>>

\section{Servo}
\label{sec:servo}
A crucial test of Servo is performance --- Servo must be at least as fast
as other browsers at similar tasks to succeed, even if it provides additional memory safety.
\tblref{servo-perf} shows a preliminary comparison of the performance of the layout stage (described
in \secref{sec:browsers}) of rendering several web sites in Mozilla Firefox's Gecko engine compared
to Servo, taken on a modern MacBook Pro.
\begin{table}
  \begin{center}
    \begin{tabular}{r || r | r | r}
      Site & Gecko & Servo 1 thread & Servo 4 threads \\
      \hline
      Reddit & 250 & 100 & 55  \\
      CNN & 105 & 50 & 35 \\
    \end{tabular}%
  \end{center}%
  \caption{Performance of Servo against Mozilla's Gecko rendering engine on the layout portion of some common sites.
  Times are in milliseconds, where lower numbers are better.}
  \label{servo-perf}
\end{table}

In the remainder of this section, we cover specific areas of Servo's design or implementation that make use of
Rust and the impacts and limitations of these features.

@@

\subsection{Compilation strategy}
Many statically typed implementations of polymorphic languages such as Standard ML of New Jersey~\cite{SMLNJ} and
\ocaml{}~\cite{ocaml-manual-3.0} have used a compilation strategy that optimizes representations of data types when
polymorphic code is monomorphically used, but defaults to a less efficient style otherwise, in order to share
code~\cite{ocaml-repr}.
This strategy reduces code size, but leads to unpredictable performance and code, as changes to a codebase that
either add a new instantiation of a polymorphic function at a given type or, in a modular compilation setting, that
expose a polymorphic function externally, can change the performance of code that is not local to the change being
made.

Monomorphization, as in MLton~\cite{mlton-compiler}, instead instantiates each polymorphic code block at each of the types
it is applied against, providing predictable output code to developers at the cost of some code duplication.
This is the strategy used by virtually all \Cplusplus{} compilers to implement templates, so it is proven and well-known
within systems programming. Rust follows this approach, although it improves on the ergonomics of \Cplusplus{} templates
by embedding serialized generic function ASTs within ``compiled'' binaries.

Rust also chooses a fairly large default compilation unit size. A Rust \textit{crate} is subject to whole-program
compilation~\cite{weeks:whole-program-mlton}, and is optimized as a unit. A crate may comprise hundreds of modules,
which provide namespacing and abstraction. Module dependencies within a crate are allowed to be cyclic.  The distinction
between crates and modules reflects the fact that computers have grown by a factor of approximately $10^9$ since C was
designed for the PDP-7, while the human capacity to understand code as a unit has regretfully not kept pace.

The large compilation unit size slows down compilation, and especially diminishes the ability to build code in parallel. It has however enabled us to write Rust code that easily matches the sequential speed of its \Cplusplus{} analog, without requiring the Servo developers to become compiler experts. Servo contains about 600 modules within 20 crates.
